# The Big (Bayesian) Picture

**Learning objectives:**

- Learn to think like a Bayesian.
- Explore the foundations of a Bayesian data analysis and how they contrast with the frequentist alternative
- Learn a little bit about the history of the Bayesian philosophy

## Thinking like a Bayesian 1/4

```{r, fig.cap="A Bayesian knowledge-building diagram"}
DiagrammeR::grViz("
  digraph thinking_bayesian{
  
  # node statement
  node [shape = oval]
  a [label = 'Prior'];
  b [label = 'Data'];
  c [label = 'Posterior'];
  d [label = 'New data'];
  e [label = 'Posterior'];
  f [label = 'New data']
  g [style = invisible ]
  
  # edge statement
  a -> c b -> c
  c -> e d -> e
  f-> g [style = dashed] e-> g [style = dashed]
  }")
```
**Both Bayesian and frequentist share a common goal: learn from data about the world around** 

Both use data to fit nodels, make predictions and evaluate hypothesis

## Quiz time! 

4-5: frequentist  
6-8: a bit of both  
9-12: Bayesian  

TODO link to script and data in repo

## Thinking like a Bayesian 2/4

###  Interpreting probability: 

- Bayesian philosophy: relative plausibility of an event 

- Frequentist philosophy: long-run relative frequency of a repeatable  event

## Thinking like a Bayesian 3/4

### Bayesian balancing act

Two claims: 

- Zuofo claims he can predict the outcome of coin flip

- Kavya claims she can distinguish between natural and artificial sweeteners 

If both succeed with a 10/10 sucess rate what can we conclude from this?

The frequentist approach will discard prior knowledge (it is harder to predict coin flip that having a sensitive palate to sweeteners) and the Bayesian want to use this prior knowledge. 

-> How can we balance Prior and Data?

## Thinking like  a Bayesian 4/4

### Asking question

What's the chance that I actually have the disease (a)?  Versus I do not have the disease, What's the chance that I would have gotten this positive test results (b)? 

```{r}
# building data
disease <- c(rep("disease", 4), rep("no disease", 96))
a <- "test positive" ; b <- "test negative"
test <- c(rep(a, 3), b, rep(a, 9), rep(b, 87))
disease_status <- data.frame(disease, test)
# contingency table
contingency_disease <- table(disease_status)
contingency_disease <- addmargins(contingency_disease)
knitr::kable(contingency_disease )
```

(a): `3 / 12`

(b): `9 / 96`

Analogy between (b) and *p-value*: it is more natural to study the uncertainty of a yet-unproven hypothesis than the uncertainty of data we have already observed.(authors'opinion) 

## Quick history lesson 

From stigmatized to being used in modeling COVID-19 rates. Why?

- advances in computing 
- departure from tradition (what people learn is what people use)
- reevaluation of subjectivity : frequentist is also subjective and subjectivity is not any more a dirty word.


## Look ahead 

### 4 units

1. Bayesian foundations: 5 chapters

Focus: models & distributions (conjugate family) 

2. Posterior simulations & analysis: 3 chapters

Focus: when conjugate is not an option: MCMC then posterior analysis

3. Bayesian regression & classification 

Focus: extending unit 1 **reponse variable (Y)** with **predictor variables (X)**

4. Hierarchical Bayesian models

Focus: expanding unit 3 to accomodate and harness grouped data : expanding unit 3 to accomodate and harness grouped data : expanding unit 3 to accomodate and harness grouped data 

## Summary 

- Posterior knowledge <- balancing information from data and prior knowledge 

- more "waves"of data -> refine knowledge (less effect of prior)

- two analysts  will converge on the same posterior knowledge

## Resources mentioned

This is a list of resources mentioned in the first meeting!

### Other Bayesian books: 

- Richard McElreath   
      * book: https://xcelab.net/rm/statistical-rethinking/  
      * vidéo: https://github.com/rmcelreath/stat_rethinking_2022

- The "puppy" book by John K. Kruschke : https://sites.google.com/site/doingbayesiandataanalysis/

### Drawing DAG (Directed Acyclic Graph)

- Pen and paper  
- [DiagrammeR](https://rich-iannone.github.io/DiagrammeR/docs.html): for drawing diagram, uses [Graphviz](https://www.graphviz.org/documentation/) or [mermaid](https://mermaid-js.github.io/mermaid/#/)  
- [Dagitty](http://dagitty.net/): for causal diagrams

### Podcast 

- [Learning Bayesian Statistics](https://podcasts.apple.com/us/podcast/how-to-teach-and-learn-bayesian-stats-with-mine-dogucu/id1483485062?i=1000526699407) ep. 42 With Mine Dogucu




## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
