# The Big (Bayesian) Picture

**Learning objectives:**

- Learn to think like a Bayesian.
- Explore the foundations of a Bayesian data analysis and how they contrast with the frequentist alternative
- Learn a little bit about the history of the Bayesian philosophy

## Thinking like a Bayesian 1/4

```{r, fig.cap="A Bayesian knowledge-building diagram"}
DiagrammeR::grViz("
  digraph thinking_bayesian{
  
  # node statement
  node [shape = oval]
  a [label = 'Prior'];
  b [label = 'Data'];
  c [label = 'Posterior'];
  d [label = 'New data'];
  e [label = 'Posterior'];
  f [label = 'New data']
  g [style = invisible ]
  
  # edge statement
  a -> c b -> c
  c -> e d -> e
  f-> g [style = dashed] e-> g [style = dashed]
  }")
```

> In Bayesian analysis, the prior represents what you know before seeing the data. The posterior then represents what you know having seen the data.

**Both Bayesian and frequentist share a common goal: learn from data about the world around. Both use data to fit nodels, make predictions and evaluate hypothesis** 



## Quiz time! 

4-5: frequentist  
6-8: a bit of both  
9-12: Bayesian  

TODO link to script and data in repo

## Thinking like a Bayesian 2/4

###  Interpreting probability: 

- Bayesian philosophy: relative plausibility of an event

- Frequentist philosophy: long-run relative frequency of a repeatable  event

## Thinking like a Bayesian 3/4

### Bayesian balancing act

Two claims: 

- Zuofo claims he can predict the outcome of coin flip

- Kavya claims she can distinguish between natural and artificial sweeteners 

If both succeed with a 10/10 sucess rate what can we conclude from this?

The frequentist approach will discard prior knowledge (it is harder to predict coin flip that having a sensitive palate to sweeteners) and the Bayesian want to use this prior knowledge. 

-> How can we balance Prior and Data?

## Thinking like  a Bayesian 4/4

### Asking question

What's the chance that I actually have the disease (a)?  Versus I do not have the disease, What's the chance that I would have gotten this positive test results (b)? 

```{r}
# building data
disease <- c(rep("disease", 4), rep("no disease", 96))
a <- "test positive" ; b <- "test negative"
test <- c(rep(a, 3), b, rep(a, 9), rep(b, 87))
disease_status <- data.frame(disease, test)
# contingency table
contingency_disease <- table(disease_status)
contingency_disease <- addmargins(contingency_disease)
knitr::kable(contingency_disease )
```

(a): `3 / 12`

(b): `9 / 96`

Analogy between (b) and *p-value*: it is more natural to study the uncertainty of a yet-unproven hypothesis than the uncertainty of data we have already observed.(authors'opinion) 

## Quick history lesson 

From stigmatized to being used in modeling COVID-19 rates. Why?

- advances in computing 
- departure from tradition (what people learn is what people use)
- reevaluation of subjectivity : frequentist is also subjective and subjectivity is not any more a dirty word.


## Look ahead 

### 4 units

1. Bayesian foundations: 5 chapters

Focus: models & distributions (conjugate family) 

2. Posterior simulations & analysis: 3 chapters

Focus: when conjugate is not an option: MCMC then posterior analysis

3. Bayesian regression & classification 

Focus: extending unit 1 **reponse variable (Y)** with **predictor variables (X)**

4. Hierarchical Bayesian models

Focus: expanding unit 3 to accomodate and harness grouped data.

## Summary 

- Posterior knowledge <- balancing information from data and prior knowledge 

- More "waves"of data -> refine knowledge (less effect of prior)

- With more and more data, two analysts will converge on the same posterior knowledge

## Resources mentioned

This is a list of resources mentioned in the first meeting!

### Other Bayesian books: 

- Richard McElreath   
      * book: https://xcelab.net/rm/statistical-rethinking/  
      * vidéo: https://github.com/rmcelreath/stat_rethinking_2022

- The "puppy" book by John K. Kruschke : https://sites.google.com/site/doingbayesiandataanalysis/

- [Introduction to Bayesian Thinking, Clyde, Centinkaya-Rundel et al](https://statswithr.github.io/book/index.html) similar level to our book

- [Bayesian Data Analysis, Andrew Gelman](http://www.stat.columbia.edu/~gelman/book/BDA3.pdf) more precise (and mathematical)

-	[Intro to Bayes Theorem, Wrath of Math, video](https://www.youtube.com/watch?v=3sJn5-Cjm2s)
	Clear explanation of the meaning of `given` in statements like probability
	of A given B. $ P(A | B) $


### Drawing DAG (Directed Acyclic Graph)

- Pen and paper  
- [DiagrammeR](https://rich-iannone.github.io/DiagrammeR/docs.html): for drawing diagram, uses [Graphviz](https://www.graphviz.org/documentation/) or [mermaid](https://mermaid-js.github.io/mermaid/#/)  
- [Dagitty](http://dagitty.net/): for causal diagrams

### Podcast 

- [Learning Bayesian Statistics](https://podcasts.apple.com/us/podcast/how-to-teach-and-learn-bayesian-stats-with-mine-dogucu/id1483485062?i=1000526699407) ep. 42 With Mine Dogucu




## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/EwK5BatQgpw")`

<details>
<summary> Meeting chat log </summary>

```
00:06:01	Olivier:	hello !
00:06:47	Olivier:	various links : https://r4ds.github.io/bookclub-bayes_rules/
00:06:55	Olivier:	https://docs.google.com/spreadsheets/d/18IDSOU2bfkD55kOB18qCB7Idbpiyp4_9qeWjkvE-Syc/edit#gid=0
00:09:44	Olivier:	Letś wait that everyone configure zoom :P
00:10:18	Gabby Palomo:	I didn't see if anyone else signed up for this cohort. I imagine more people did, right?
00:10:46	Olivier:	I do not know the number of participant
00:10:52	Olivier:	I will have to ask 
00:11:08	Will:	Well hopefully. I'm assuming for fellow Brits there are lots people celebrating the jubilee.
00:11:59	Gabby Palomo:	ah that's true!!
00:12:10	erik.aalto@tocaboca.com:	I’m not hearing anything, but I could hear recording in progress when I joined..
00:12:24	Olivier:	do you hear me ? 
00:12:27	Gabby Palomo:	There are 85 people in the slack channel so will see.
00:12:42	erik.aalto@tocaboca.com:	Nope, cant hear anything:/
00:13:15	Will:	Yes I think I can hear Erik
00:13:45	erik.aalto@tocaboca.com:	Darn, can’t hear anything…this is weird
00:13:57	Olivier:	we hear you at least
00:14:03	erik.aalto@tocaboca.com:	I hear the zoom notifs…like ”recording in progress”
00:15:52	Olivier:	it is fine 
00:16:13	Olivier:	is it better
00:17:24	Olivier:	is it good now ?
00:17:41	Olivier:	i restaart it
00:19:02	Olivier:	working or not ? 
00:19:20	Ronald Legere:	Not yet… there is a audio test thing in the audio settings menu
00:19:48	Olivier:	good name 
00:56:58	Ronald Legere:	Can you link those podcasts to slack? Or here ;)
01:06:31	Olivier Leroy:	DiagrammeR
```
</details>


### Cohort 2

`r knitr::include_url("https://www.youtube.com/embed/GvK98TXNb18")`


### Cohort 3

`r knitr::include_url("https://www.youtube.com/embed/Vy9ii3nCics")`

<details>
<summary> Meeting chat log </summary>

```
00:18:42	Mary EMERAGHI:	Hello Everyone... So, I will drop a brief on myself here since I am in a different timezone. I am a beginner with little or no knowledge of this subject save that I have knowledge of Basic Statistics and would want to advance deeper to use it for my plant breeding work on Rice and Cereals
00:19:05	Mary EMERAGHI:	My Name though is Mary Emeraghi from Nigeria
00:19:16	Mary EMERAGHI:	Currently in Cote d'Ivoire
00:20:59	Mary EMERAGHI:	Could I get the full details of these books especially on Theoretical knowledge to help me
00:21:01	Mary EMERAGHI:	?
00:21:47	Mary EMERAGHI:	thanks Gabby
00:21:54	Mary EMERAGHI:	By who please?
00:21:55	Gabby Palomo:	Bayesian Models: A statistical primer for ecologists; Hobbs and Hooten
00:22:21	Gabby Palomo:	Doing Bayesian Data Analysis: A tutorial with R, JAGS, and Stan. John Kruschke.
00:22:56	Mary EMERAGHI:	much appreciated
00:47:48	Mary EMERAGHI:	no
00:47:55	Mary EMERAGHI:	sorry
01:05:06	Oluwafemi Oyedele:	Thank you !!!
```
</details>


### Cohort 4

`r knitr::include_url("https://www.youtube.com/embed/KdnnwyGGst4")`

<details>
<summary> Meeting chat log </summary>

```
00:01:05	Derek Sollberger (he/him):	textbook: https://www.bayesrulesbook.com/
00:01:13	Derek Sollberger (he/him):	cohort notes: https://r4ds.github.io/bookclub-bayes_rules/index.html
00:01:20	Derek Sollberger (he/him):	schedule: https://docs.google.com/spreadsheets/d/1tf2GwxwpZNyQsXfyi8wBfuMyRrr6FT1ZxZAQPzxFJCA/edit#gid=0
```
</details>
