# Bayes’ Rule

**Learning objectives:**

- Explore foundational probability tools
    - conditional probability: Probability of A given B $P(A|B)$  
    - joint probability: Probability of A and B occuring  $P(A \cap B)$   
    - marginal probability:  Probability of an event $P(A)$
    - Law of Total Probability: If a probability of an event is unknown it can be calculated using the known probability of other related event   
    
- Conduct first formal Bayesian analysis  

- Practice your Bayesian grammar  
    - Prior
    - Likelihood
    - Normalizing constant  
    
- Simulate Bayesian models
    - `sample()`
    - `sample_n()`
    - `rbinon()`

## Building a Bayesian model for events 

**First Data set** `?? fake_news` 

```{r, fig.cap="Bayesian knowledge-building diagram for wether or not the article is fake", echo=FALSE}
DiagrammeR::grViz("
  digraph {
  
  # node statement
  node [shape = oval]
  a [label = 'Prior: 40% of the article are fake'];
  b [label = 'Data: ! more common among fake news'];
  c [label = 'Posterior: Is the article fake or not?!'];

  # edge statement
  a -> c b -> c
  }")
```

Two variables:  
- fake or real  
- presence of ! or not  

### Workflow: 

1. Prior probability model  
2. A model for interpreting the data  
3. Posterior probability model  

$$ P(FakeNew = 0.4 )  \quad and \quad  P(Real = 0.6) $$
$$ P(B = 0.4 )  \quad and \quad  P(B^c = 0.6) $$

Here $P(FakeNew)$ : prior probability of an article to be a fake news

Valid probability model :   
1. accounts all event    
2. assign probabilities for each event  
3. sum to one  

### Conditional probability & likelihood

$P(ExClam)$ : probability that an article contains an exclamation mark in his title

We know that if an article is fake news : 26.67% that the title contains **!** and if it is not fake this is just 2.22%.

$$ P(Exclam|FakeNew) = 0.2667 \quad and \quad P(Exclam|Real = 0.0222)  $$

This is a **conditional probability**. Conditional probability help know if B give us insight in A. If it does not provide any information it means that A and B event are **independent** ($P(A|B) = P(A)$). 

If we get an new article with `!` in his title is it more *likely* to be fake or real news? Here we are checking that $P(Exclam|FakeNew) >>  P(Exclam|Real = 0.0222)$. We will call this procedure a **likelihood function** it allow us to evaluate the relative compatibility of data with event $B$ or $B^c$ (Real of Fake). 


### Normalizing constant 

Then we need can calculate the **joint probability** (probability of observing A and B for example), of each options (here it is half of them): 

$$P(Exclam \cap FakeNew) = P(Exclam|FakeNew) P (Fakenew) = 0.2667 *0.4 = 0.1067$$ 

ie: the joint probability of observing an exclamation point in the title in a fake news is equal of the probability of having an exclamation point given that the article is fake times the probability of the article to be fake. 


$$P(PasExclam \cap FakeNew ) = (1 - P(Exclam|FakeNew)) * P(FakeNew)) = (1 - 0.2667) * 0.4 = 0.2993$$ 

Here $P(B), P(B^c), P(A), P(A^c)$ cover the **marginal distributions**, they are the sum "marges" of the table. Inside of the contingency table we have the **joint probibilties**.  $P(Fake)$ will be the **marginal probability** of "Fake". 

```{r}
Fake <- c(0.1067, 0.2933, 0.4)
Real <- c(0.0133, 0.5867, 0.6)
Total <- c(0.12, 0.88, 1)
joint_p <- data.frame(Fake, Real, Total, row.names = c("Exclam", "No Exclam", "Total"))
knitr::kable(joint_p)
```


We can also summarize it like that : 

![](image/proba-1.vsg)


### Posterior probability via Bayes' Rule!

$P(Exclam)$ is our normalizing constant.

Ok but we want $P(FakeNew|exlam)$ ie $P(B|A)$

$$P(FakeNew|exclam) = \frac{P(exclam \cap FakeNew)}{P(Exclam)} = \frac{P(FakeNew)L(FakeNew|Exclam)}{P(Exclam)}$$ 

$$posterior = \frac{prior . likelihood}{normalizing \quad constant} = \frac{0.4 * 0.2667}{0.12} = 0.889$$

## Posterior simulation 

This was a model. Now we will use a simulation! 

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
set.seed(84735)
# Define possible articles
article <- data.frame(type = c("real", "fake"))

# Define the prior model
prior <- c(0.6, 0.4)
article_sim <- dplyr::sample_n(article, size = 10000, 
                        weight = prior, 
                        replace = TRUE)
# dats model
article_sim <- article_sim %>% 
  mutate(data_model = case_when(type == "fake" ~ 0.2667,
                                type == "real" ~ 0.0222))


# Simulate exclamation point usage 
data <- c("NoExclam", "Exclam")

set.seed(3)
# Rbase simplier ? 
article_sim <- article_sim %>%
  group_by(1:n()) %>% 
  mutate(usage = sample(data, size = 1, 
                        prob = c(1 - data_model, data_model)))

ggplot(article_sim, aes(x = type)) + 
  geom_bar() + 
  facet_wrap(~ usage)

```

## Example Pop vs Soda vs Coke

 Expend TRUE/FALSE example with one with categories. 
 
## Building a Bayesian model for random variables (1/n)

### First step prior

$\pi$ : skill of Kasparov relative to Deep Blue (**random variable**) 

Prior model of $\pi$:

```{r}
pi <- c(0.2, 0.5, 0.8, "total")
# pmf : probability mass functions
pmf <- c(0.1, 0.25, 0.65, 1)
prior_pi <- data.frame(pi, pmf)
knitr::kable(t(prior_pi))
```

## Building a Bayesian model for random variables (1/n)

### Binomial data model

Y is the number of games (on 6 games) that Kasparov wins.

Y our random variable: {0, 1, ...., 6}  , depends on $\pi$  

$$f(y|\pi) = P (Y = y|\pi )  $$
$y$ : any possible outcone

If we assume games are independents (no effect on each other) and $\pi$ is fixed we can use the *Binomial model*.

Y is the number of successes in a fixed number of trials ($n$)

$$ Y|\pi \sim Bin(n, \pi) $$ 

$$ f(y|\pi) = \begin{pmatrix} 6 \\y \end{pmatrix} \pi^y(1 - \pi)^{6 - y} \quad for \quad  y \in \begin{Bmatrix} 0, 1, 2, 3, 4, 5, 6 \end{Bmatrix}  $$ 
We can use the prior for $\pi$ and all $y$ to calculate each probabilities. 

## Building a Bayesian model for random variables (1/n)

### Binomial likelihood function

Kasparov only won one of six games. This is our data with $L(\pi|y = 1)$. We can calculate it for each $\pi$ value. 

What is more likely is that $pi$ was 0.2.

### Probability mass functions vs likelihood functions

When $\pi$ is known the conditional pmf allows us to compare the probabilities of different possible value of data Y (y1, y2 ..) occuring with $pi$

when Y = y is known the likelihood function allows us to to compare the relative values of $\pi$ ($\pi_1, \pi_2, etc ..$)

### Normalizing constant

Total probability that Kasparov would win Y = 1 game across all possible win probability of $\pi$

We apply the Law of Total Probability (the sum of all likelihood for each value of $\pi$ by the prior probailities of these $\pi$ values)

$$ f(y = 1) = L(\pi = 0.2 | y = 1) f(\pi = 0.2 ) +  L(\pi = 0.5 | y = 1) f(\pi = 0.5 ) + L(\pi = 0.8 | y = 1) f(\pi = 0.8 ) +$$ 

$$ f( y = 1) \simeq 0.3932 * 0.1 + 0.0938 * 0.25 + 0.0015 * 0.65 \simeq 0.0637 $$

### Posterior probability model 

We have the prior, the likelihod and the normalizing constant -> Bayes Rules!


$$ posterior = \frac{prior . likelihood}{normalizing \quad constant} \quad for \in  \begin{Bmatrix} 0.2, 0
5, 0.8 \end{Bmatrix} $$


### Posterior shortcut

The normalizing constant, is a **constant** it appears in all posterior calculations. We can work with unnormalizied proabilities or we can divide each unnormalizied probability by the sum of each of them.

$$ posterior \propto prior * likelihood $$
## Posterior simulation 

```{r}
# Define possible win probabilities
chess <- data.frame(pi = c(0.2, 0.5, 0.8))

# Define the prior model
prior <- c(0.10, 0.25, 0.65)

# Simulate 10000 values of pi from the prior
set.seed(84735)
chess_sim <- sample_n(chess, size = 10000, weight = prior, replace = TRUE)

chess_sim <- chess_sim %>% 
  mutate(y = rbinom(10000, size = 6, prob = pi))

# Focus on simulations with y = 1
win_one <- chess_sim %>% 
  filter(y == 1)


# Plot the posterior approximation
ggplot(win_one, aes(x = pi)) + 
  geom_bar()

```

## Links shared in the second meeting

Bayesian probability for babies! (with cookies) : https://raw.githubusercontent.com/epimath/epid-814-materials/master/Lectures/BayesianEstimation.pdf
Author: Marisa Eisenberg


## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/-6f60j96d58")`

<details>
<summary> Meeting chat log </summary>

```
00:11:44	erik.aalto@tocaboca.com:	Hello, sorry for joining late, happy to see you again 😄
00:36:06	Brendan Lam:	https://www.youtube.com/watch?v=HZGCoVF3YvM
00:37:26	Erik:	we’ll utilize the following likelihood function notation  
<poorly formatted paste>
00:37:41	Erik:	sorry for how the formatting above went…
00:42:42	Lisa:	to add to brendan's resource, here is a slide deck using cookies as an example: https://github.com/epimath/epid-814-materials/blob/master/Lectures/BayesianEstimation.pdf
01:02:00	Federica Gazzelloni:	thanks
01:02:34	Federica Gazzelloni:	I’d like to go back to the L(..|..)
01:02:35	Gabby Palomo:	It was clear Olivier. Don’t worry!!
01:03:07	Federica Gazzelloni:	just to understand how it works and what is the difference
01:03:43	Erik:	Need to jump out! Thanks for today.
01:04:15	Lisa:	I need to get back to work too. thank you for today!
01:04:58	Federica Gazzelloni:	thanks
01:05:03	Federica Gazzelloni:	I’ll check that
```
</details>

### Cohort 2

`r knitr::include_url("https://www.youtube.com/embed/2gP7rLLLUsA")`

`r knitr::include_url("https://www.youtube.com/embed/QirQynPol0k")`
